---
title: "Modeling and prediction for movies"
author: "Raymond Ngiam Heng Aik"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(GGally)
library(leaps)
```

### Load data

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

**How the observations in the sample are collected:**

The data set is comprised of 651 randomly sampled movies produced and released before 2016.
<br></br>

**Implications of this data collection method on the scope of inference (generalizability):**

Since the data collection involves random sampling, the sample can be considered representative of the population. Hence findings from this observational study can be generalized to the entire population.
<br></br>

**Implications of this data collection method on the scope of inference (causality):**

Since this is an observational study, not an experiment where subjects are randomly assigned for specific treatments, hence, the findings on relationship among different factors only establish association but do not imply causality.

* * *

## Part 2: Research question

**Do relationship exist among Rotten Tomatoes' audience score and the rest of the variables in the "movies.Rdata" dataset?**

This research question is of interest because based on past historical records we can get insights on which variables are having correlation to Rotten Tomatoes' audience score, which is related to the popularity of movies among audiences. We can then build a linear model to predict the popularity of future movies on average, based on combination of the correlated variables. 

* * *

## Part 3: Exploratory data analysis

<span style="text-decoration:underline">Variables to consider for the full model</span>

1. title_type

2. genre

3. runtime

4. mpaa_rating

5. imdb_rating

6. imdb_num_votes

7. critics_scores

8. critics_rating

9. best_pic_nom

10. best_pic_win

11. best_actor_win

12. best_actress_win

13. best_dir_win

14. top200_box

<span style="text-decoration:underline">Excluded variables:</span>

1. title

2. studio

3. thtr_rel_year

4. thtr_rel_month

5. thtr_rel_day

6. dvd_rel_year

7. dvd_rel_month

8. dvd_rel_day

9. director

10-14. actor1-actor5

15. imdb_url

16. rt_url

<span style="text-decoration:underline">Reasons for exclusion variables:</span>

Since we are interested to study properties of movies that are influential to the movie's popularity in general, the variable "title", "studio", "director", and "Actor1"-"Actor5" are excluded because they are too specific. If we use these variables in creating the regression model, the model will only be reasonable in the context of the existing sample of movies. We will not be able to make predictions for movies other that those with "title" among the existing sample. Same goes for "studio", "director", and "Actor1"-"Actor5".

Time related variables such as theatre release date and DVD release date are also excluded because we are interested in long term popularity of movies, not popularity during release in theatre or release of DVD.

Finally, IMDB url and Rotten Tomatoes url are also excluded as they are obviously irrelevant to context of this problem.

```{r}
movies_no_na <- na.omit(movies)
N = nrow(movies_no_na)
N
```

**Univariate analysis**

One of the numeric explanatory variables that is of particular interest is **imdb_num_votes**.

Below is the histogram for the distribution of **imdb_num_votes**,

```{r}
ggplot(data = movies_no_na, aes(x=imdb_num_votes)) +
  geom_histogram(binwidth = 1500) +
  labs(title = "Histogram of imdb_num_votes")
range(movies_no_na$imdb_num_votes)
```

The range of **imdb_num_votes** is (183, 893008), and it is heavily right-skewed. It will be more meaningful if we apply a log transformation, and name it as a new variable, **limdb_num_votes**.

```{r}
movies_added_predictors <- movies_no_na %>%
  mutate(limdb_num_votes = log10(imdb_num_votes))
```

Histogram of **limdb_num_votes**
```{r}
ggplot(data = movies_added_predictors, aes(x=limdb_num_votes)) +
  geom_histogram(binwidth = 0.1) +
  labs(title = "Histogram of limdb_num_votes")
```

<br> </br>

**Bivariate analysis between explanatory variable and response variable:**

1) Title type and audience score

Box plot to visualize the differences in audience score distribution among categorical variable, title type.
```{r}
ggplot(data=movies_added_predictors, aes(x=title_type, y=audience_score)) + 
  geom_boxplot() + 
  labs(title = "Boxplot of Audience Score against Title Type")
```

Summary statistics
```{r}
movies_added_predictors %>%
  group_by(title_type) %>%
  summarize(mean=mean(audience_score), sd=sd(audience_score),median=median(audience_score),IQR=IQR(audience_score),min=min(audience_score),max=max(audience_score), count=n())
```

<span style="color:blue">Interpretation: There are differences in the center of the distribution of audience score among different classes of title type in this data sample. The "Documentary" title type has the highest audience score on average compared to the rest.</span>
<br> </br>

2) Genre and audience score

Box plot to visualize the differences in audience score distribution among categorical variable, genre.
```{r}
ggplot(data=movies_added_predictors, aes(x=genre, y=audience_score)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Boxplot of Audience Score against Genre")
```

Summary statistics
```{r}
movies_added_predictors %>%
  group_by(genre) %>%
  summarize(mean=mean(audience_score), sd=sd(audience_score),median=median(audience_score),IQR=IQR(audience_score),min=min(audience_score),max=max(audience_score), count=n())
```

<span style="color:blue">Interpretation: There are differences in the center of the distribution of audience score among different classes of genre in this data sample. The "Documentary" genre has the highest audience score on average compared to the rest.</span>
<br> </br>

3) Critics rating and audience score

Box plot to visualize the differences in audience score distribution among categorical variable, genre.
```{r}
ggplot(data=movies_added_predictors, aes(x=critics_rating, y=audience_score)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Boxplot of Audience Score against Critics Rating")
```

Summary statistics
```{r}
movies_added_predictors %>%
  group_by(critics_rating) %>%
  summarize(mean=mean(audience_score), sd=sd(audience_score),median=median(audience_score),IQR=IQR(audience_score),min=min(audience_score),max=max(audience_score), count=n())
```

<span style="color:blue">Interpretation: There are differences in the center of the distribution of audience score among different classes of critics rating in this data sample. The "Certified Fresh" critics rating has the highest audience score on average compared to the rest.</span>
<br> </br>

4) Runtime and audience score

Scatter plot to visualize the correlation between runtime and audience score.
```{r}
ggplot(data=movies_added_predictors, aes(x=runtime, y=audience_score)) + 
  geom_point() + 
  geom_smooth(method="lm", se=FALSE) + 
  labs(title = "Scatter plot of Audience Score against Runtime")
```

Summary statistic on the correlation.
```{r}
cor(movies_added_predictors$runtime, movies_added_predictors$audience_score)
```

<span style="color:blue">Interpretation: There is a weak, positive correlation of 0.20 between runtime and audience score.</span>
<br> </br>

5) IMDb Rating and audience score

Scatter plot to visualize the correlation between IMDb rating and audience score.
```{r}
ggplot(data=movies_added_predictors, aes(x=imdb_rating, y=audience_score)) + 
  geom_jitter() + 
  geom_smooth(method="lm", se=FALSE) + 
  labs(title = "Scatter plot of Audience Score against IMDb Rating")
```

Summary statistic on the correlation.
```{r}
cor(movies_added_predictors$imdb_rating, movies_added_predictors$audience_score)
```

<span style="color:blue">Interpretation: There is a strong positive correlation of 0.86 between IMDb rating and audience score.</span>
<br> </br>

6) Critics Score and audience score

Scatter plot to visualize the correlation between critics score and audience score.
```{r}
ggplot(data=movies_added_predictors, aes(x=critics_score, y=audience_score)) + 
  geom_jitter() + 
  geom_smooth(method="lm", se=FALSE) + 
  labs(title = "Scatter plot of Audience Score against Critics Score")
```

Summary statistic on the correlation.
```{r}
cor(movies_added_predictors$critics_score, movies_added_predictors$audience_score)
```

<span style="color:blue">Interpretation: There is a strong positive correlation of 0.70 between critics score and audience score.</span>

<br> </br>

**Collinearity analysis among numeric explanatory variables:**

Next, we will explore the correlation among the numeric explanatory variables.

```{r}
ggpairs(dplyr::select(movies_added_predictors,runtime,limdb_num_votes,imdb_rating,critics_score),title = "Pairwise relationships among numeric explanatory variables")
```

<br> </br>

We observed that **imdb_rating** and **critics_score** are highly correlated with correlation coefficient of 0.762, while the rest are not as correlated. Since the two variables are measure of the movie's popularity, we can compute a weighted average from the two variables to form a new variable **score**. 

Since **imdb_rating** is of the range (0,10) and **critics_score** is of the range (0,100), we reweight them by multiply **imdb_rating** with weight 10, and **critics_score** with weight 1, then taking the average of both. The final **score** variable will be of the range (0,100).

```{r}
movies_added_predictors <- movies_added_predictors %>%
  mutate(score = (imdb_rating*10+critics_score)/2)
```

* * *

## Part 4: Modeling

**Model parameters**

Continue from our explanatory data analysis, we ended up with **13** explanatory variables as listed below. We will use these updated explanatory variables as the full model for our model selection: 

+ **title_type**
+ **genre**
+ **runtime**
+ **mpaa_rating**
+ <span style="text-decoration:line-through">imdb_rating</span> 
<span style="color:blue;font-weight:bold">score</span>
+ <span style="text-decoration:line-through">imdb_num_votes</span> 
<span style="color:blue;font-weight:bold">limdb_num_votes</span>
+ <span style="text-decoration:line-through">critics_score</span>
+ **critics_rating**
+ **best_pic_nom**
+ **best_pic_win**
+ **best_actor_win**
+ **best_actress_win**
+ **best_dir_win**
+ **top200_box**

<br> </br>

**Model selection**

The baseline linear model with all 13 explanatory variables has an adjusted R squared value of **0.6623**.

```{r}
data_train<-movies_added_predictors %>%
  dplyr::select(audience_score,title_type,genre,runtime,score,limdb_num_votes,critics_rating, best_pic_nom,best_pic_win,best_actor_win,best_actress_win,best_dir_win,top200_box)
modelFull = lm(audience_score ~ ., data=data_train)
modelFullSummary<-summary(modelFull)
modelFullSummary$adj.r.squared
```

For model selection, we use adjusted R-squared as the model selection criteria. Adjusted R-squared is chosen over p-value because it produces more reliable prediction. We will use the *regsubsets* function from the **leaps** package to produce sets of candidate models with different sizes of variable.

```{r}
leaps<-regsubsets(audience_score~.,data=data_train,nbest=1,nvmax=13,method="exhaustive")
leaps
plot(leaps,scale="adjr2",main="Adjusted R^2")
```

We then take the candidate model of specific number of variable with the highest adjusted R squared. This normally corresponds to the candidate model largest number of variables as it has more chance to improve the explained variability and also contains the rest of the variables.

```{r}
leaps.summary<-summary(leaps)
plot(leaps.summary$adjr2, xlab = "num of Variables", ylab = "AdjR^2", main = "AdjR^2 vs num of variables")
idxAdjRSqrd<-which.max(leaps.summary$adjr2)
points(idxAdjRSqrd, leaps.summary$adjr2[idxAdjRSqrd], pch = 20, col = "red")
```

This maybe otherwise for other model selection criteria like the BIC (we minimize the BIC instead of maximizing it):

```{r}
plot(leaps.summary$bic, xlab = "num of Variables", ylab = "BIC", main = "BIC vs num of variables")
idxBICMin<-which.min(leaps.summary$bic)
points(idxBICMin, leaps.summary$bic[idxBICMin], pch = 20, col = "red")
```

Next, we determine the variables that are included in the model with the highest adjusted R squared, and then create the final model, **movies_mlr_final**.

```{r}
coef(leaps,idxAdjRSqrd)
movies_mlr_final<-lm(audience_score ~ genre + score + limdb_num_votes + critics_rating + best_pic_nom + best_actress_win, data=data_train)
movies_mlr_final
```

<br> </br>

**Model diagnostics**

<span style="text-decoration:underline">Linearity:</span>

Plot of residuals vs fitted values
```{r}
ggplot(data = movies_mlr_final, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals") +
  labs(title = "Residuals vs Fitted")
```

<span style="color:blue">Random scatter around zero in residual plot. Hence relationship between response variable and explanatory variables should be linear.</span>

<span style="text-decoration:underline">Nearly normal residuals:</span>

Histogram of residuals

```{r}
plot(density(movies_mlr_final$residuals))
```

Normal probability plot of residuals

```{r}
qqnorm(movies_mlr_final$residuals)  
qqline(movies_mlr_final$residuals) 
```

<span style="color:blue">From the histogram, residual distribution of the current model is slightly skewed to the right. And from the normal probability plot, there are some deviation from normality at the higher end. The condition for nearly normal residuals is only fairly satisfied.</span>

<span style="text-decoration:underline">Constant variability:</span>

<span style="color:blue">From the residuals vs fitted plot, the condition for constant variability is also only fairly satisfied.</span>

<br> </br>

**Interpretation of model coefficients**
```{r}
summary(movies_mlr_final)
```

<span style="text-decoration:underline">Intercept:</span>

Movies in the **Action & Adventure genre**, with **critics rating Certified Fresh**, **no Oscar best picture nomination**, **no Oscar best actress**, **zero log of IMDb number of votes**, and **zero score** are expected on average to have audience score of **-24.37**.

<span style="text-decoration:underline">Slopes:</span>

All else held constant, the model predicts that the Animation genre have **8.33** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Art House & International genre have **11.60** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Comedy genre have **2.36** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Documentary genre have **12.94** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Drama genre have **3.60** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Horror genre have **4.52** point **lower** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Musical & Performing Arts genre have **11.87** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Mystery & Suspense genre have **3.81** point **lower** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Other genre have **4.93** point **higher** in audience score than the Action & Adventure genre on average.

All else held constant, the model predicts that the Science Fiction & Fantasy genre have **1.42** point **lower** in audience score than the Action & Adventure genre on average.

All else held constant, for each 1 point increase in score the model predicts the audience score to be **higher** on average by **0.93** point.

All else held constant, for each 1 point increase in log of IMDb number of votes the model predicts the audience score to be **higher** on average by **5.12** point.

All else held constant, the model predicts that the critics rating Fresh have **1.64** point **higher** in audience score than the the critics rating Certified Fresh on average.

All else held constant, the model predicts that the critics rating Rotten have **9.10** point **higher** in audience score than the the critics rating Certified Fresh on average.

All else held constant, the model predicts that movies with Oscar best picture nomination have **4.13** point **higher** in audience score on average than the ones without Oscar Best Picture Nomination.

All else held constant, the model predicts that movies with no Oscar best actress have **2.71** point **lower** in audience score on average than the ones which did not win.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

