---
title: "Modeling and prediction for movies"
author: "Raymond Ngiam Heng Aik"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

**How the observations in the sample are collected:**

The data set is comprised of 651 randomly sampled movies produced and released before 2016.
<br></br>

**Implications of this data collection method on the scope of inference (generalizability):**

Since the data collection involves random sampling, the sample can be considered representative of the population. Hence findings from this observational study can be generalized to the entire population.
<br></br>

**Implications of this data collection method on the scope of inference (causality):**

Since this is an observational study, not an experiment where subjects are randomly assigned for specific treatments, hence, the findings on relationship among different factors only establish association but do not imply causality.

* * *

## Part 2: Research question



* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

