---
title: "Modeling and prediction for movies"
author: "Raymond Ngiam Heng Aik"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("movies.Rdata")
```



* * *

## Part 1: Data

**How the observations in the sample are collected:**

The data set is comprised of 651 randomly sampled movies produced and released before 2016.
<br></br>

**Implications of this data collection method on the scope of inference (generalizability):**

Since the data collection involves random sampling, the sample can be considered representative of the population. Hence findings from this observational study can be generalized to the entire population.
<br></br>

**Implications of this data collection method on the scope of inference (causality):**

Since this is an observational study, not an experiment where subjects are randomly assigned for specific treatments, hence, the findings on relationship among different factors only establish association but do not imply causality.

* * *

## Part 2: Research question

**Do relationship exist among Rotten Tomatoes' audience score and the rest of the variables in the "movies.Rdata" dataset?**

This research question is of interest because based on past historical records we can get insights on which variables are having correlation to Rotten Tomatoes' audience score, which is related to the popularity of movies among audiences. We can then build a linear model to predict the popularity of future movies on average, based on combination of the correlated variables. 

* * *

## Part 3: Exploratory data analysis

<span style="text-decoration:underline">Variables to consider for the full model</span>

1. title_type

2. genre

3. runtime

4. mpaa_rating

5. imdb_rating

6. imdb_num_votes

7. critics_scores

8. critics_rating

9. best_pic_nom

10. best_pic_win

11. best_actor_win

12. best_actress_win

13. best_dir_win

14. top200_box

<span style="text-decoration:underline">Excluded variables:</span>

1. title

2. studio

3. thtr_rel_year

4. thtr_rel_month

5. thtr_rel_day

6. dvd_rel_year

7. dvd_rel_month

8. dvd_rel_day

9. director

10-14. actor1-actor5

15. imdb_url

16. rt_url

<span style="text-decoration:underline">Reasons for exclusion variables:</span>

Since we are interested to study properties of movies that are influential to the movie's popularity in general, the variable "title", "studio", "director", and "Actor1"-"Actor5" are excluded because they are too specific. If we use these variables in creating the regression model, the model will only be reasonable in the context of the existing sample of movies. We will not be able to make predictions for movies other that those with "title" among the existing sample. Same goes for "studio", "director", and "Actor1"-"Actor5".

Time related variables such as theatre release date and DVD release date are also excluded because we are interested in long term popularity of movies, not popularity during release in theatre or release of DVD.

Finally, IMDB url and Rotten Tomatoes url are also excluded as they are obviously irrelevant to context of this problem.

* * *

## Part 4: Modeling

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 5: Prediction

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button above. Make sure that your code is visible in the project you submit. 
Delete this note when before you submit your work.

* * *

## Part 6: Conclusion

